Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-22 ICML-submission/introduction.tex
tweaked to fit into 8 pages

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between the learner and the adversary. At each
timestep $t$, the adversary chooses a labeled example $(x_t, y_t)$ where the
feature vector $x_t \in \R^d$ is revealed to the learner and the correct label
$y_t$ is kept secret by the adversary. Upon receiving the feature vector $x_t$,
the learner is asked to make a prediction $\widehat{y}_t$. Immediately after
making its prediction, the learner receives feedback. In contrast with the
standard full-information setting (where the feedback given is the correct label
$y_t$), here the feedback is only a binary indicator of whether the prediction
was correct or not. The protocol of the problem is formally stated below.

@@

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definition~\ref{definition:weak-linear-separability}.
The standard notion of multiclass linear
separability~\citep{Crammer-Singer-2003} corresponds to the weak linear
separability. For binary classification, the weak and strong separability are
identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, weak linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

In the full-information feedback setting, it is well known
\citep{Crammer-Singer-2003} that if all examples have norm at most $R$ and are
weakly linearly separable with a margin $\gamma$, then the \textsc{Multiclass
Perceptron} algorithm makes at most $\lfloor 2(R/\gamma)^2 \rfloor$ mistakes.
%%%%%
It is also known that any (possibly randomized) algorithm must
make $\frac{1}{2} \left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the
worst case.
These results guarantee that the \textsc{Multiclass Perceptron} achieves an
information-theoretically optimal mistake upper bound, meanwhile it has
low time and memory complexities.%
%The guarantees provided by \textsc{Multiclass Perceptron} is very satisfying,
%since it achieves an information-theoretically optimal mistake upper bound,
%while having low time and memory complexities.
%%%%%
~\footnote{
%%%%%
%These are folklore results. 
For completeness, we present these 
folklore
%%%%%
results along
with their proofs in Appendix~\ref{section:multiclass-perceptron-proofs} in the
supplementary material.}

>>>>

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between a learner and an adversary. At each
timestep $t$, the adversary chooses a labeled example $(x_t, y_t)$ and reveals $x_t\in \R^d$ to the learner.  
Upon receiving $x_t$, the learner makes a prediction $\widehat{y}_t$ and receives feedback. 
In contrast to the
standard full-information setting, where the feedback given is the correct label
$y_t$, here the feedback is only a binary indicator of whether the prediction
was correct or not. The protocol of the problem is formally stated below.

@@

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definition~\ref{definition:weak-linear-separability}.
The standard notion of multiclass linear
separability~\citep{Crammer-Singer-2003} corresponds to the weak linear
separability. %For binary classification, the weak and strong separability are
%identical. However, they differ for multiclass classification with $K \ge 3$
%classes. 
For multiclass classification with $K$ classes, weak linear
separability requires that all examples from the same class lie in an intersection of $K-1$ halfspaces and all other examples lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

In the full-information feedback setting, it is well known
\citep{Crammer-Singer-2003} that if all examples have norm at most $R$ and are
weakly linearly separable with a margin $\gamma$, then the \textsc{Multiclass
Perceptron} algorithm makes at most $\lfloor 2(R/\gamma)^2 \rfloor$ mistakes.
%%%%%
It is also known that any (possibly randomized) algorithm must
make $\frac{1}{2} \left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the
worst case.
The \textsc{Multiclass Perceptron} achieves an
information-theoretically optimal mistake bound, while being
computationally and memory efficient.
~\footnote{
%%%%%
%These are folklore results. 
For completeness, we present these 
folklore
%%%%%
results along
with their proofs in Appendix~\ref{section:multiclass-perceptron-proofs} in the
supplementary material.}
